---
title: "Heart Disease Visualization"
author: "Mingwei Wu"
output: pdf_document
---

#Background
## The database contains 76 attributes, but all published experiments refer to using a subset of 14 of them, at particular, the Cleveland database is 

##Data Description
###age: patients' age
###sex: 1 is male, 2 is female
###cp: chest pain type 4 level
###trestbps:  resting blood presure
###chol: serum cholestoral in mg/dl
###fbs: fasting blood sugar >120 mg/dl is value 1
###restecg: resting electrocardiographic reults (values 0,1,2)
###thalach: maximum heart rate achieved
###exang: exercise incuced angina
###oldpeak: ST depression induced by exercise relative to rest
###slope: the slope of the peak exercise ST segment
###ca: number of major vessels (0-3) colored by flourosopy
###thal: 3 is normal; 6 fixed defect; 7 reversable defect
###target: 1 has heart disease, 0 not
```{r}
data<-read.csv("heart.csv",header = TRUE)
head(data)
```
```{r}
pairs(data) #paris data to see the relationship in numeric values
```


```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
heart<-data%>%
  mutate(gender= ifelse(sex==1,"Male","Female"),
         chest_pain_level= ifelse(cp==0,"normal",
                                  ifelse(cp==1,"mild",
                                         ifelse(cp==2,"moderate","severe"))),
         fblood_sugar=ifelse(fbs==1,">120","<=120"),
         rest_electrocardigoraphic= ifelse(restecg==0,"normal",
                                           ifelse(restecg==1,"abnormalily","definite")),
         exercise=ifelse(exang==1,"yes","no"),
        heart_condition=ifelse(target==1,"yes","no")) # rebuild the column to the data frame
```

```{r}
heart%>%
  group_by(gender)%>%
  summarise(gender_rate=mean(target)) # calculate the heart disease  rate of the gender
```
```{r}
heart%>%
  ggplot(aes(heart_condition,fill=gender))+geom_bar()+ggtitle("gender vs heart condition") #data visualization to display the relationship
```

```{r}
heart%>%
  group_by(exercise)%>%
  summarise(exercise_rate=mean(target))
```


```{r}
heart%>%
  ggplot(aes(heart_condition,fill=exercise))+geom_bar()+ggtitle("exericse vs heart condition")
```
```{r}
heart%>%
  group_by(age)%>%
  count()%>%
  filter(n>5)%>%
  ggplot(aes(age,n))+geom_col()+ggtitle("age analysis") # checking a proportion of heart condition with age
```

```{r}
heart%>%
  ggplot(aes(heart_condition,fill=fblood_sugar))+geom_bar()+ggtitle("blood sugar > 120 vs heart condition")
```
```{r}
heart%>%
  ggplot(aes(x=age,y=chol,color=heart_condition))+geom_point()+geom_smooth(method="lm")+ggtitle("relationship age and chol vs heart_condition in gender")+ facet_wrap(~gender) #plot the point and line for data visualization
```
```{r}
heart%>%
  ggplot(aes(gender,trestbps))+geom_boxplot()+xlab("Sex")+ylab("resting blood pressure")+facet_grid(~chest_pain_level)
```












