---
title: "Heart Disease Visualization"
author: "Mingwei Wu"
output: pdf_document
---

# Background
## The database contains 76 attributes, but all published experiments refer to using a subset of 14 of them, at particular, the Cleveland database is the only one that has been used by ML researchers to this data.
## Data Description
### age: patients' age
### sex: 1 is male, 2 is female
### cp: chest pain type 4 level
### trestbps:  resting blood presure
### chol: serum cholestoral in mg/dl
### fbs: fasting blood sugar >120 mg/dl is value 1
### restecg: resting electrocardiographic reults (values 0,1,2)
### thalach: maximum heart rate achieved
### exang: exercise incuced angina
### oldpeak: ST depression induced by exercise relative to rest
### slope: the slope of the peak exercise ST segment
### ca: number of major vessels (0-3) colored by flourosopy
### thal: 3 is normal; 6 fixed defect; 7 reversable defect
### target: 1 has heart disease, 0 not

## Import data
```{r}
data<-read.csv("heart.csv",header = TRUE)
head(data)
```
### First step, we compare the relationship on every column. Double check the dependent value.
```{r}
pairs(data) #pairs data to see the relationship in numeric values
```

### Edit the column with categories. Change the int value to character values.
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
heart<-data%>%
  mutate(gender= ifelse(sex==1,"Male","Female"),
         chest_pain_level= ifelse(cp==0,"normal",
                                  ifelse(cp==1,"mild",
                                         ifelse(cp==2,"moderate","severe"))),
         fblood_sugar=ifelse(fbs==1,">120","<=120"),
         rest_electrocardigoraphic= ifelse(restecg==0,"normal",
                                           ifelse(restecg==1,"abnormalily","definite")),
         exercise=ifelse(exang==1,"yes","no"),
        heart_condition=ifelse(target==1,"yes","no")) # rebuild the column to the data frame
```

## Geder Analysis
### Calculate the rate of heart disease in th gender. The attached result below, the rate of Female in database has 75% who had heart disease, and the Male rate had approximate 45%
```{r}
heart%>%
  group_by(gender)%>%
  summarise(gender_rate=mean(target)) # calculate the heart disease  rate of the gender
```
## Question: why female has higher proportion in heart disease.
### Count the quantity of gender. In the dataframe, the male quantity is domain, almost double quantity than female.
```{r}
a<-table(heart$gender)
par(mfrow=c(1,2))
barplot(a,
        col=rainbow(2),
        legend=rownames(a),
        xlab="Gender",
        ylab="Count",
        main="Barplot of Gender")
pct<-round(a/sum(a)*100)
lbs<-paste(c("Female","Male"), "",pct,"%",sep="")
pie(a,labels=lbs, main="Percentage of Gender") #data visualization to compare quantity of gender
```
### Barplot displays the disease quantity of gender that are almost same. on the contract, the Male count higher than female.
```{r}
b<-table(heart$gender,heart$heart_condition)
barplot(b,
        col = c("red","blue"),
        legend=rownames(b),
        beside=TRUE,
        xlab="sick",
        ylab="count",
        main="side by side barplot") # data visualization to compare count with disease.
```

## Exercise Analysis
### The rate of exercise with disease is 23%, and the rate of no_exercise with disease approximate 70%
```{r}
heart%>%
  group_by(exercise)%>%
  summarise(exercise_rate=mean(target))
```
### Only 33% in the dataframe who do exercise everyday.
```{r}
a1<-table(heart$exercise)
par(mfrow=c(1,2))
barplot(a1,
        col=rainbow(2),
        legend=rownames(a1),
        xlab="exercise",
        ylab="Count",
        main="Barplot of exercise")
pct<-round(a1/sum(a1)*100)
lbs<-paste(c("no","yes"), "",pct,"%",sep="")
pie(a1,labels=lbs, main="Percentage of exericise") #data visualization to compare quantity of gender
```
### The proportion of people with no disease who do not exercise are almost same, but the proportion of disease, no-exercise is extreme higher than do-exercise.
```{r}
b1<-table(heart$exercise,heart$heart_condition)
barplot(b1,
        col = c("red","blue"),
        legend=rownames(b1),
        beside=TRUE,
        xlab="disease",
        ylab="count",
        main="side by side barplot") # data visualization to compare count with disease.
```
## Age Analysis
### the age histogram shows the normal distribution, and estimate the high proportion of disease of age range 50-60.
```{r}
par(mfrow=c(1,2))
hist(heart$age,labels=TRUE,main="Histogram of Age",xlab = "age",ylab = "frequency")
boxplot(heart$age,horizontal = TRUE,col="red",main="boxplot of age")
```
```{r}
qqnorm(heart$age,frame=FALSE)
qqline(heart$age,col="steelblue",lwd=2)
```

## Fasting Blood Sugar Analysis
### The fasting blood sugar either less than 120 or greater than 120, it is not significat effect to the disease.
```{r}
heart%>%
  group_by(fblood_sugar)%>%
  summarise(fblood_sugar_rate=mean(target))
```

```{r}
heart%>%
  ggplot(aes(heart_condition,fill=fblood_sugar))+geom_bar()+ggtitle("blood sugar > 120 vs heart condition")
```
## Cholestoral Analysis
### Plot the dependent value of chol and age to disease. For the Female, age and chol with disease display increased trend line, but non-disease is decreased. However, for the male, either disease or not, both trends increased line.
```{r}
heart%>%
  ggplot(aes(x=age,y=chol,color=heart_condition))+geom_point()+geom_smooth(method="lm")+ggtitle("relationship age and chol vs heart_condition in gender")+ facet_wrap(~gender) #plot the point and line for data visualization
```

## Resting electrocardiographic Analysis
```{r}
heart%>%
  ggplot(aes(gender,trestbps))+geom_boxplot()+xlab("Sex")+ylab("resting blood pressure")+facet_grid(~chest_pain_level)
```
## Maximum Heart Rate Achieved Analysis
### No matter with gender, if people have disease, the maximum heart rate achieved is higher than non-disease, also decresed with age. the non-disease heart rate achieved is relatively stable.
```{r}
heart%>%
  ggplot(aes(age,thalach,color=heart_condition))+geom_point()+geom_smooth(se=FALSE)+facet_grid(~gender)+ggtitle("maxium heart rate vs gender and target")
```

## Principle Component Analysis
```{r}
pca<-prcomp(heart[,4:10],scale=TRUE)
pca
```
```{r}
library(factoextra)
screeplot(pca)
fviz_screeplot(pca) #plot the tendency of principle component analysis
```
```{r}
pca$sdev^2
pca$rotation
```
```{r}
fviz_pca(pca) #data visualization for pca
```

## Statistics Analysis
### Randomly split the data into 70% train_set and 30% test_set for logistic regression.
```{r}
library(caTools)
set.seed(927)
sample<-sample.split(data$target,SplitRatio=0.70)
train_set<-subset(data,sample==TRUE)
test_set<-subset(data,sample==FALSE)
head(heart,10)
```
```{r}
logistic<-glm(target~.,train_set,
              family=binomial())
summary(logistic)
```
## Remove insignificant factors
### After removing the insignificant factors, the AIC value is going down.
```{r}
logistic1<-glm(target~sex+cp+thalach+oldpeak+ca+thal,
               train_set,
               family=binomial())
summary(logistic1)
```
### coefficients build up formular
```{r}
logistic1$coefficients
```
### prediction on test_set
```{r}
pred<-predict(logistic1,test_set,type="response")
pred_new<-as.data.frame(pred)
categorise<-function(x){
  return(ifelse(x>0.5,1,0))
}
pred_new<-apply(pred_new,2,categorise)
head(pred_new,10)
```
## Model Evaluation
### Model has 85.7% accuracy in predecting future data with logistic regression model.
```{r}
library(caret)
confusionMatrix(as.factor(test_set$target),as.factor(pred_new))
```
